paper2_resistance_mechanisms_mixtures

### 2017-01-06 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
bibliography: IR.bib
csl: biomed-central.csl
---

@@

(This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)

## ANDY TO DO
1. add a screenshot of the UI
1. add to intro including about about results in levick et al.
1. explain resistance-restoration and its relation to selection coefficient
1. expand discussion to include relevance to real world 
1. in discussion address potential criticism that these results are just for a limited region of paramter space (a. point to Levick et al. that explored more of space, b. point to UI that allows readers to explore spce for themselves and identify whether there are places where this does not hold).
1. ? briefly refer to linkage disequilibrium and say we are investigating.

1. can I develop explanatory diagrams to describe the input parameters ?
1. mention other potential effects e.g. costs of resistance, sex linkage, ...
1. add to discussion about relevance of these mechanisms in the field e.g. exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation. (Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
1. mention in discussion that repellency would lead to decrease in exposure
1. is there a refernce to say that the effectiveness of insecticides can be modified at the design stage ?

@@

Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potentially huge consequences for malaria burden. To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones. There are different potential strategies for using new insecticides and recomendations for how best to preserve their utility for controlling mosquitoes and malaria. There is however limited recent, accessible modelling work to support suggestions of which strategies may be best in different situations. Previous work suggests that preferred strategies will be situation specific. This points to a need for an accessible, mechanistic understanding of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy. Such a mechanistic understanding would be more robust to uncertainty in parameter values in field situations.

Here we use a relatively simple model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures. We target a readership with limited experience of population genetics and resistance. The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies. We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide,  and how much the resistance counteracts the effect of the insecticide.

We show firstly that using a second insecticide, at the same time as a first, always slows evolution of resistance to the first. Secondly we show that the ability of an insecticide to kill the susceptible mosquitoes (effectiveness), is the most important factor influencing whether using insecticides in a sequence or a mixture is likely to lead to faster resistance to both of them. These results are explained in terms of the mechanisms that cause them. We discuss how these results can inform intervention strategies in the field.

We describe an R package and user-interface that allows further exploration of the development of resistance. The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.

@@

Insecticide resistance is a problem for malaria [@Who2012][@Ranson2016] other vector borne diseases and agriculture.

@@

We use a 2-locus population genetic model of insecticide resistance [@Levick2017] to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies. Levick et al. (in press) describe the technical details of the model and explore large parameter ranges. We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance. This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide use strategies.

@@

The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual. One locus is represented for each insecticide. There are two potential alleles per locus. Each allele confers either resistance or susceptibility to the insecticide. The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model. These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).

@@

Here we principally look at model outputs of the change in allele frequencies over time (generations). We investigate three insecticide use strategies.

1. single insecticide
1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches 50%
1. two insecticides used in a mixture (concurrently)

The model is implemented in R [RCoreTeam2016], the code is hosted on Github.

@@

### Single insecticide
For single insecticide use, faster resistance spread resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 2-4). In the plots each input was varied in isolation with the remaining set to 0.5 (except for starting frequency of resistance which was set to 0.01 in all bar Fig 4).

@@

When single insecticide use was represented (Figs 2-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%. Higher values of the five inputs investigated all led to more rapid increases in resistance.

For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 2A to 2B). This makes sense as vector kill is effectively a product of exposure times effectiveness. For example exposing 50% of a  population to an insecticide which is 75% effective would be expected to have the same result as exposing 75% of the population to one which is 50% effective. This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance. In both cases the increased deaths of susceptible vectors cause a higher selection pressure that can explain the faster increase.

The pattern of more rapid increase in resistance at higher dominance levels (Fig 3A) can be explained by increased survival of heterozygotes in the presence of the insecticide. Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival. Thus selection pressure for the resistance allele will be increased because it confers more of an advantage when only present on one chromosome.

The faster development of resistance under higher resistance-restoration (Fig 3B) can be explained by it's effect on the survival of the resistant genotypes. Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.

@@

@@

### TODO ADD TO DISCUSSION : link these refs to the results here
[@Mani1985]
"use of mixtures is always more effective in delaying the onset of resistance, often by many orders of magnitude"
~ assumed that SS always killed i.e. effectiveness=1
~ exposure fixed at 0.9

[@Gould1986]
~ insecticidal toxins genetically engineered into plants
~ additive rtaher than multiplicative fitness
~ effectiveness ~0.2-0.5 (might be expected to favour sequences by our results)
~ no clearly superior strategy

[@Roush1998]
~ identified mortality of SS homozygotes (our “insecticide effectiveness” parameter) as critical (his Fig 3) and concluded (page 1784) :
“As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective”

[@Georghiou1983]
potential IRM strategies :
~ moderation : preserve susceptible genes by limiting selection pressure
~ saturation : high dose so that heterozygous resistants are killed
~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance

@@

```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Effectiveness, B. Exposure"}

@@

  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
         theme_bw() +  
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) +  
         labs(colour = "effectiveness") +
         coord_trans(y = "log10") +    
         geom_line()

       )

  exposure <- 0.2
  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')

@@

  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
         theme_bw() +
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) +
         labs(colour = "exposure", y=NULL) +
         coord_trans(y = "log10") +   
         geom_line()

       )

```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Dominance, B. Resistance restoration"}

  #single insecticide dominance fig

@@

  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
         theme_bw() +
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) + 
         labs(colour = "dominance") +
         coord_trans(y = "log10") + 
         geom_line()

       )

  # single insecticide resistance restoration fig.

@@

  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
         theme_bw() +
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) +
         labs(colour = "resistance restoration", y=NULL) +
         coord_trans(y = "log10") +   
         geom_line()   
       )

@@

```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of resistance starting frequency on single insecticide use."}

@@

```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="*KEY FIG OF PAPER* Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences. A. All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B. Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C. Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D. Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence"}

@@

```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Higher effectivenesses for both insecticides, times-to-resistance longer for the mixture in all scenarios. A. Effectiveness of insecticide1 increased from 0.5 to 0.8, B. Effectiveness of insecticide1 increased from 0.5 to 1, C. Effectiveness of insecticides 1 and 2 increased to 0.8, D. Effectiveness of insecticide2 0.8 and of insecticide1 1"}

@@

```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A. All control inputs equal at 0.5, B. Resistance restoration for insecticide1 increased from 0.5 to 0.8, C. Dominance for insecticide1 increased to 0.8, D. Resistance restoration and dominance for insecticide1 increased to 0.8"}

@@

```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A. All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B. Effectiveness for insecticide1 increased from 0.5 to 0.8, C. Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D. Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.001 to 0.0001"}

@@

```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A. All control inputs equal at 0.5, male exposure 0.8, B. Effectiveness for insecticide1 increased from 0.5 to 0.8, C. Male exposure decreased from 0.8 to 0.2, D. Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}

>>>>

---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
bibliography: library.bib
csl: biomed-central.csl
#header-includes:
#    - \usepackage{setspace}
#    - \doublespacing
#    - \usepackage{lineno}
#    - \linenumbers
---

```{r, eval=FALSE, include=FALSE}
#!!!! paper protocol
#!!!! might be easier to put text & figs into separate Rmd files after draft stage
#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
#!!!! submission figs, set this chunk to eval=TRUE

# options to create final publication quality figures
# causes the document to fail, but the figs do get stored in pap2figs/
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='pap2figs/')
```

@@

## ANDY TO DO
1. mention single gene vs additive effects (with ref.)
1. ? briefly refer to linkage disequilibrium and say we are investigating OR include a figure OR refer to one of my existing figures
1. can I develop explanatory diagrams to describe the input parameters ?
1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage ?

@@

_must be 350 words, currently 375 with subheadings Background, Results & Conclusions_

##### Background
Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potentially huge consequences for malaria burden. To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones. There are different potential strategies for using new insecticides and recomendations for how best to preserve their utility for controlling mosquitoes and malaria. There is however limited recent, accessible modelling work to support suggestions of which strategies may be best in different situations. Previous work suggests that preferred strategies will be situation specific. This points to a need for an accessible, mechanistic understanding of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy. Such a mechanistic understanding would be more robust to uncertainty in parameter values in field situations.

Here we use a relatively simple model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures. The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies. We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide,  and how much the resistance counteracts the effect of the insecticide.

##### Results
We show firstly that using a second insecticide, at the same time as a first, always slows evolution of resistance to the first. Secondly we show that the ability of an insecticide to kill the susceptible mosquitoes (effectiveness), is the most important factor influencing whether using insecticides in a sequence or a mixture is likely to lead to faster resistance to both of them. These results are explained in terms of the mechanisms that cause them. We discuss how these results can inform intervention strategies in the field.

##### Conclusions
We describe an R package and user-interface that allows further exploration of the development of resistance. The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.

## Keywords
insecticide resistance; public health; mosquitoes; vector-borne diseases; infectious diseases; malaria; population genetics

@@

Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].

Public health efforts in controlling malaria have centred on treated bed nets and indoor residual spraying, both reliant on insecticides. Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million malaria cases between 2000 and 2015 [@Bhatt2016].

A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.

The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016]. Several new insecticides are now in development [@IVCC2016] and very soon decisions will need to be made about how best to use them. The model presented here can help inform strategies to minimise the developm,ent of resiostance in new and existing insecticides.

@@

_TODO add in here about theory of IR & protocols for resistance management._

We use a 2-locus population genetic model of insecticide resistance [@Levick2017] to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies. In an earlier paper [@Levick2017] we described the technical details of the model and explore large parameter ranges. We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance. This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Rousch1989][@IRAC2011][@FAO2012][@GPIRM2012].

@@

The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual. One locus is represented for each insecticide. There are two potential alleles per locus. Each allele confers either resistance or susceptibility to the insecticide. Fitness is the main currency of the model, representing how much each genotype survives and reproduces, with a maximum value of 1. The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model. These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1). Each of these inputs can be varied as a proportion between 0 and 1. We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide. The more conventional 'selection coefficient' is calculated by multiplying Resistance restoration by insecticide effectiveness. Doing it this way allows us to vary all inputs between 0 and 1 without fitness going above 1 in the model.

@@

Here we principally look at model outputs of the change in resistance allele frequencies over time (generations). Resistance allele frequencies are calculated from the frequencies of each genotype as follows.

> Resistance allele frequency = 2*RR + RS

We investigate three insecticide use strategies.

1. single insecticide
1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches a threshold of 50%
1. two insecticides used in a mixture (concurrently)

The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an online user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.

@@

### Single insecticide
For single insecticide use, higher values of insecticide effectiveness, exposure, dominance or resistance restoration resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach the resistance threshold of either 0.5 or 0.75.

Similarly, higher values of starting resistance frequency also resulted in faster spread of resistance (Fig 4) and shorter times to resistance thresholds.

@@

When single insecticide use was represented (Figs 3-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%. Higher values of the five inputs investigated all led to more rapid increases in resistance.

For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B). This makes sense as vector kill is effectively a product of exposure times effectiveness. For example exposing 50% of a  population to an insecticide which is 75% effective would be expected to have the same result as exposing 75% of the population to one which is 50% effective. This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance. In both cases the increased deaths of susceptible vectors cause a higher selection pressure that can explain the faster increase.

The pattern of more rapid increase in resistance at higher dominance levels (Fig 3C) can be explained by increased survival of heterozygotes in the presence of the insecticide. Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival. Thus selection pressure for the resistance allele will be increased because it confers more of an advantage when only present on one chromosome.

The faster development of resistance under higher resistance-restoration (Fig 3D) can be explained by it's effect on the survival of the resistant genotypes. Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.

@@

It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria. As such our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide. Using no insecticide could have serious consequences for mosquito and malaria abundance. The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while and sustaining that in the longer term by restricting the development of resistance. Others are looking at combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].

The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ransom2016]. The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of paramter space (e.g. holding most inputs constant at intermediate values while modifying single inputs in isolation). Thus there is the potential that the system may behave differently in the field than we describe here. However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017]. In addition we provide an online user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown here.

#### discussion: putting results into context (some parts may want to go earlier)
I will now put these predicted responses of resistance into the context of previous work and recomendations.

Most of these predicted reponses of resistance are consistent with recomendations that have been produced by WHO[@WHO2012], IRAC[@IRAC2011] and FAO[@FAO2012].

[Next bit might want to go in intro, but not if I refer to our results]
Earlier modelling work mostly from the 1980s has made sometimes contrasting predictions and these can be explained by the details of their structure and the particular situations they represent. One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985]. That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide. This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9. As described earlier we particular predict that the very high effectiveness is likely to favour mixtures over sequences.

[@Roush1989]

The model we describe is more flexible and as such can reproduce similar results to these earlier approaches and be used to represent more detailed, operationally relevant strategies.

Relevance of these mechanisms in the field e.g. exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation. [@Birget2015a] mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
_TODO check which of the Birget papers mention this_

One consistent aspect of the results presented here can at first seem to go against current advice on good practice. We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5). This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance. In contrast the widely held view seems to be that low effectiveness or exposure associated with poor implementation can promote the development of resistance. e.g. from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance. These include: ... • the use of application rates that are below or above those recommended on the label; • poor coverage of the area being treated ...". The explanation for this apparent difference is in the effect of dominance. Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?]. Our model does show that increasing dominance leads to faster spread of resaistance (Fig 3C, 7). So, the effect of poor implementation of an insecticide intervention on resistance could be eiether positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure. Our model cobined with laboratory and field results could help to assess these trade-offs.

[@Taillebois2016]
Laboratory experiments on aphids showing that in low dose mixtures insecticides most often had additive effects as represented in our model, but that also commonly mortality was higher than expected (synergistic) which we can also represent within the model.

[@Roush1989]
~ identified mortality of SS homozygotes (our “insecticide effectiveness” parameter) as critical (his Fig 3) and concluded (page 1784) :
“As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective”

[@Georghiou1994]
potential IRM strategies :
~ moderation : preserve susceptible genes by limiting selection pressure
~ saturation : high dose so that heterozygous resistants are killed
~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance

@@

_todo mention other potential effects e.g. costs of resistance, sex linkage, ..._

_TODO : Should I try to cite in press Killeen about combining interventions ?_

TODO FINAL CONCLUDING PARAGRAPH
~ Effectiveness and exposure are key and can be modified.
~ The modelling approach can be used to investigate more detailed strategies.

@@

![Screenshot of one online model user interface, accessible at : https://andysouth.shinyapps.io/resistmob2/](paper2_UI1.png)

```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on changing resistance frequency over time with single insecticide use. A. Effectiveness, B. Exposure, C. Dominance, D. Resistance restoration. For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01). "}

  library(cowplot)

  #list to put plots in for cowplot
  plotlist <- list(4)

  #A effectiveness  
  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')

@@

  #A  
  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
         theme_bw() +  
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) +  
         labs(colour = "effectiveness") +
         coord_trans(y = "log10") +    
         geom_line()

  #B exposure
  exposure <- 0.2
  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')

@@

  #B  
  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
         theme_bw() +
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) +
         labs(colour = "exposure", y=NULL) +
         coord_trans(y = "log10") +   
         geom_line()

  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')

  #C single insecticide dominance
  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')

@@

  #C 
  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
         theme_bw() +
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) + 
         labs(colour = "dominance") +
         coord_trans(y = "log10") + 
         geom_line()

  #D single insecticide resistance restoration

@@

  #D  
  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
         theme_bw() +
         theme(legend.position = "bottom", legend.key = element_blank()) + 
         guides(colour = guide_legend(reverse=TRUE)) +
         labs(colour = "resistance restoration", y=NULL) +
         coord_trans(y = "log10") +   
         geom_line()

  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')

@@

```{r Fig3, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of changing resistance starting frequency on changing resistance frequency over time with single insecticide use."}

@@

```{r Fig4, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="*KEY FIG OF PAPER* Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences. A. All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B. Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C. Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D. Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence"}

@@

```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Higher effectivenesses for both insecticides, times-to-resistance longer for the mixture in all scenarios. A. Effectiveness of insecticide1 increased from 0.5 to 0.8, B. Effectiveness of insecticide1 increased from 0.5 to 1, C. Effectiveness of insecticides 1 and 2 increased to 0.8, D. Effectiveness of insecticide2 0.8 and of insecticide1 1"}

@@

```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A. All control inputs equal at 0.5, B. Resistance restoration for insecticide1 increased from 0.5 to 0.8, C. Dominance for insecticide1 increased to 0.8, D. Resistance restoration and dominance for insecticide1 increased to 0.8"}

@@

```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A. All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B. Effectiveness for insecticide1 increased from 0.5 to 0.8, C. Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D. Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.001 to 0.0001"}

@@

```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A. All control inputs equal at 0.5, male exposure 0.8, B. Effectiveness for insecticide1 increased from 0.5 to 0.8, C. Male exposure decreased from 0.8 to 0.2, D. Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
